---
title: "Lab 2 - Permutation?"
author: "Your Group Member Names Here!"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE}
library(infer)
library(tidyverse)
library(viridis)
```

# Grading the Professor

Many college courses conclude by giving students the opportunity to evaluate the course and the instructor anonymously. However, the use of these student evaluations as an indicator of course quality and teaching effectiveness is often criticized because these measures may reflect the influence of non-teaching related characteristics, such as the physical appearance of the instructor. The article titled, "Beauty in the classroom: instructors' pulchritude and putative pedagogical productivity" (Hamermesh and Parker, 2005) found that instructors who are viewed to be better looking receive higher instructional ratings. (Daniel S. Hamermesh, Amy Parker, Beauty in the classroom: instructors pulchritude and putative pedagogical productivity, Economics of Education Review, Volume 24, Issue 4, August 2005, Pages 369-376, ISSN 0272-7757, 10.1016/j.econedurev.2004.07.013. http://www.sciencedirect.com/science/article/pii/S0272775704001165.)

In this lab we will analyze the data from this study in order to learn what goes into a positive professor evaluation.


# Data for Today's Lab  

The data are gathered from end of semester student evaluations for a large sample of professors from the University of Texas at Austin. These data are merged with descriptors of the professors and the classes. In addition, six students rated the professors' physical appearance. The result is a data frame where each row contains a different course and each column has information on either the course or the professor. [Data are available through the Open Intro Statitics course webpage](https://www.openintro.org/stat/data/evals.php).

This is a real dataset that has been used in a publication, *Looking Good on Course Evaluations*, published in 2013 in CHANCE volume 26 issue 2.  

__1. Load the data into your work space and inspect it to answer the following questions:__   

* __How many observations (rows) do the data have?__  
<!-- Your answer here!  -->



* __How many variables (columns) do the data have?__ 
<!-- Your answer here!  -->



* __What class (data types) are the variables in the dataset?__  
<!-- Your answer here!  -->



```{r data, message = FALSE, warning = FALSE}
# read_csv does not coerce strings to factors
evals <- read_csv("data/evals.csv")

glimpse(evals)

```

</br>
</br>

__2. Is this an observational study or an experiment? The original research question posed in the paper is whether being a man leads directly to the differences in course evaluations.__ 
__(a) Given the study design, is it possible to answer this question as it is phrased?__ 
<!-- Your answer here!  -->



__(b) If not, rephrase the research question that can be addressed.__ 
<!-- Your answer here!  -->



## Cleaning the Data 

Unfortunately, the individuals responsible for recording the data incorrectly labeled the variable `gender`. Run the code below to change the name of this variable to its correct label of `sex`. 

```{r}
# data cleaning 
evals <- evals %>% 
          rename(sex = gender) # renames the gender variable to sex
```

__Bonus: What issues could arise with data labeled in this way? How would these options for gender impact the study participants or people reading the article's results?__  
<!-- Your answer here!  -->




# Plotting the Data 

First, let's create a visualization of the distribution of course evaluation scores. The code below produces a 

```{r}
ggplot(data = evals, mapping = aes(x = score)) + 
  geom_density(fill = "blue") + 
  labs(
    x = "Course Evaluation Score", 
    y = "Density"
  )

```

__3. Describe the distribution of `score`. Is the distribution skewed? What does this tell you about how students rate courses? Is this what you expected to see? Why, or why not?__
<!-- Your answer here!  -->




Next, let's compare the course evaluation scores and average beauty scores of 
the two sexes of professors. 

```{r}
# Side-by-side dot plot
ggplot(data = evals, mapping = aes(x = sex, y = score)) + 
  # defines what variables are mapped to different attributes of the plot
  
  geom_jitter(aes(color = sex)) + 
  # tells R to plot jittered dots and color them by sex
  
  scale_color_manual(values = viridis(n = 3)) +  
  # a nicer color palette than the default 
  
  labs(
    y = "Course Evaluation Score", # y-axis label
    x = "Sex of Professor (self identified)", # x-axis label
    color = "Sex" # color label
  )

```

```{r}
# side-by-side violin plots 
ggplot(data = evals, mapping = aes(x = sex, y = bty_avg)) + 
  # defines what variables are mapped to different attributes of the plot
  
  geom_violin(aes(fill = sex)) + 
  # tells R to plot violins of the density of the points for each sex 
  
  geom_jitter() + 
  # tells R to plot jittered dots on top of the violins 
  
  scale_fill_manual(values = viridis(n = 3)) +  
  # a nicer color palette than the default 
  
  labs(
    y = "Professor's Average Beauty Score", # y-axis label
    x = "Sex of Professor (self-identified)", # x-axis label
    fill = "Sex" # color label
  )

```


__4. Based on the plots above what can you say about the course evaluation scores and average beauty scores for professors who identified as male verses those who identified as female?__  
<!-- Your answer here!  -->




## Infer Package 

The `infer` package performs statistical inference using an expressive statistical grammar that coheres with the the syntax we have been using thus far in the course. 

- Each stage of the testing or estimation process is explicit and easily interpretable. 
- Stages of the process are chained together using the pipe operator (`%>%`), to be consistent with the syntax of other packages we will be using (e.g. `dplyr`).  
- Stages in the plotting process are chained together using the addition operator (`+`), to be consistent with the  `ggplot` syntax we will be using.  


## Permutation Test for a Difference in Mean Average Beauty: Example 

```{r, diffmeans}
# observed statistic
diff_mean <- evals %>% 
              specify(bty_avg ~ sex) %>%  
              # specify the model formula (y ~ x)
              calculate(stat = "diff in means",  
                        # what statistic you want calculated
                        order = c("male", "female") 
                        # what order of subtraction you want
                        )

permute <- evals %>% 
           filter(is.na(bty_avg) == FALSE) %>% 
           # taking out any courses with NA for beauty scores
           specify(bty_avg ~ sex) %>% 
           # specify the model formula 
           hypothesize(null = "independence") %>% 
           # specify the null hypothesis 
           generate(reps = 1000, type = "permute") %>%
           # generate a specified number of permutations
           calculate(stat = "diff in means", 
                     # what statistic you want calculated
                     order = c("male", "female")
                     # what order of subtraction you want
                     )
  

# p-value for hypothesis test
permute %>% # using the vector of permuted differences in means
  get_p_value(obs_stat = diff_mean, # declare the observed difference in means  
              direction = "two_sided" # state the direction for the test
              ) 


# plot of p-value
permute %>% # using the vector of permuted differences in means
  visualize(method = "simulation") + 
  # creates a histogram of the permutations
  
  shade_p_value(diff_mean, # declare the oberved difference in means
                direction = "both"
                # shades the direction specified
                )  + 
  
  labs(
    x = "Difference in Mean Average Beauty Score (Male - Female)",  
    # x-axis label
    title = "Permutation for a Difference in Means" 
    # plot title 
    )


```

## Permutation Test for a Difference in Course Evaluation Scores: Your Turn!

__5. Using the template above, create a permutation distribution of the differences in mean course evaluation scores between male and female professors.__  

```{r}
# Your code here!



```


__6. Based on the permutation test, what would you conclude for the relationship between course evaluation scores and professor's sex? Does this agree with what you saw in the side-by-side dotplot of course evaluation scores? Why or why not?__ <!-- Your answer here!  -->
 



__7. Why is the above test a permutation test and not a randomization test?__  
<!-- Your answer here!  -->




## Setting GitHub Account 

When you open your RStudio Cloud account, you have to tell RStudio how to interact with GitHub. 
To do this you will need to carry out the following steps __every time__ you open up RStudio Cloud:

1. load in the usethis package (first line of code below)  
2. specify your GitHub username and email (second line of code below)  
3. check to make sure RStudio recognizes your GitHub credentials (third line of code below)


```{r, eval = FALSE}
# To commit, push, and pull your changes to GitHub

# First -- load up the usethis package
library(usethis) 

# Second -- specify your GitHub username and email address
use_git_config(user.name = "atheobold", user.email = "allisontheobold@montana.edu")

# Third -- check by running a git situation-report
git_sitrep()
```

Once you have carried out these three steps, you can interact with GitHub through the RStudio Cloud server. Some helpful pointers for using GitHub:  

* always pull before you start working on the project (someone could have made changes while you were away!)  
* make small changes and commit and push often (this both takes less time for the computer to update your files and takes less memory to remember what changes you made)  
* __never__ work at the same time as another person (if you want to work at the same time, then work together on one computer)


# References 

[Infer Package Reference Page](https://infer.netlify.com/index.html)

[Open Intro Statistics](http://htmlpreview.github.io/?https://github.com/andrewpbray/oiLabs-base-R/blob/master/multiple_regression/multiple_regression.html)
